---
title       : Calculating the area of a triangle with Shiny
subtitle    : Developing Data Products Assignment
author      : Bela Czeiner
job         : Coursera student
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax, bootstrap]    # {mathjax, quiz, bootstrap}
ext_widgets : {rCharts: [libraries/nvd3]}
mode        : selfcontained # {standalone, selfcontained, draft}
knit        : slidify::knit2slides
---

# Calculate the area of a triangle with Shiny

The solution calcuates the are of a triangle using two methods 
and compares the results:

    1. Calculate it using Heron's formula, the classic geometry method 
    2. A geometric approximation

### Source data

The user need to enter:

    - Coordinates of the cornerns of a triangle in a Cartesian coordinate system 
    - A measure of granularity, which determines the number of data points for the the approximation

---
# Calculation with Heron's formula
Heron's formula uses the semi-perimeter (one-half the perimeter) and the measures of the three sides:

```{r, echo = F, fig.height=4, fig.width=4}
x <- c(0,2,6,0) 
y <- c(5,0,8,5)
m <- cbind(x, y)
plot(m, type="b", xlab="", ylab="")
text(1,2,"a")
text(3,7,"b")
text(4,3,"c")
```
$  Area = \sqrt{s(s - a)(s - b)(s - c)}$ ,       where $s = \frac{a + b + c}{2}$

The same formula using the coordinates of the cornerns of the triangle:
$$Area =  \frac{\left\lvert (x_1*y_2 + x_2*y_3 + x_3*y_1 - x_1*y_3 - x_2*y_1 - x_3*y_2) \right\rvert}{2}$$
(_For more details see:_ https://en.wikipedia.org/wiki/Heron%27s_formula)

---
# Calculation with geometric approximation

The second solution uses a square fitted around the triangle filled with a matrix of dots or grid points. The area of the triangle is calculated as the ratio of the number of grid points within the triangle (_TriDots_) versus the total number of grid points (_AllDots_) multiplied by the the size of the whole suare (_SqSize_).

```{r echo = F, results = 'asis', message=FALSE, fig.height=5, fig.width=5}
## prepare trianle example
x <- c(0,2,6,0); y <- c(5,0,8,5)
x <- x - min(x); y <- y - min(y)
m <- cbind(x, y)

## Create data points
gran <- 20 #granurality
maxx <- max(x); maxy <- max(y)
scaleX <- maxx/gran; scaleY <- maxy/gran 
pX <- seq(0, maxx, scaleX); #pX <- pX[c(-1,-length(pX))]
pY <- seq(0, maxy, scaleY); #pY <- pY[c(-1,-length(pY))]

grid <- data.frame(x = as.numeric(), y = as.numeric())
for (i in (1:length(pX))) {
    for (j in (1:length(pY))) {
        grid0 <- c(x = pX[i], y = pY[j])
        grid <- rbind.data.frame(grid, grid0)
    }
}
names(grid) <- c("x","y")

## Plot triangle and points inside and outside the triagle
plot(m, type="b", xlab="", ylab="")
dots.IN <- 0
for (k in (1:dim(grid)[1])) {
    side <- NULL
    for (line in (1:(length(x)-1))) {
        s0 <- ((x[line+1] - x[line]) * (grid[k,2] - y[line]) 
               - (y[line+1] - y[line]) * (grid[k,1] - x[line]))
        side <- c(side, s0)
    }
    IN_OUT <- ifelse(sum(abs(side)) == abs(sum(side)), "green", "blue") # "in", "out"
    if (IN_OUT == "green") {dots.IN <- dots.IN +1}
    points(grid[k,1], grid[k,2], pch = 19, col= IN_OUT)    
}
```
$  Area = \frac{TriDots}{AllDots}  *  SqSize$


---
# Comparison or results

```{r echo = F, results = 'asis', message=FALSE}
    # set default values
    x <- c(0,2,6,0) 
    y <- c(5,0,8,5)
    m <- cbind(x, y)

    triangle <- paste0("1st corner (x, y): ", m[1,1], ", ", m[1,2]
              , "; 2nd corner (x, y): ", m[2,1], ", ", m[2,2]
        , "; 3rd corner (x, y): ", m[3,1], ", ", m[3,2])

#     ## Triagle area with Heron's formula
#     AreaHeron <- round((abs(x[1]*y[2] + x[2]*y[3] + x[3]*y[1] 
#                         - x[1]*y[3] - x[2]*y[1] - x[3]*y[2]) / 2), 2)

    # create coordinates of grid points 
    gran <- 10
    maxx <- max(x); maxy <- max(y)
    scaleX <- maxx/gran 
    scaleY <- maxy/gran 
    pX <- seq(0, maxx, scaleX);
    pY <- seq(0, maxy, scaleY);

    grid <- data.frame(x = as.numeric(), y = as.numeric())
    for (i in (1:length(pX))) {
        for (j in (1:length(pY))) {
            grid0 <- c(x = pX[i], y = pY[j])
            grid <- rbind.data.frame(grid, grid0)
        }
    }
    names(grid) <- c("x","y")
    
    ## PLot the triangle with the coloured gridpoints
    dots.IN <- 0
    for (k in (1:dim(grid)[1])) {
        side <- NULL
        ## Point inside or outside of polygon?
        for (line in (1:(length(x)-1))) {
            s0 <- ((x[line+1] - x[line]) * (grid[k,2] - y[line]) 
                   - (y[line+1] - y[line]) * (grid[k,1] - x[line]))
            side <- c(side, s0)
        }
        IN_OUT <- ifelse(sum(abs(side)) == abs(sum(side))
                         , "green", "blue") # "in", "out"
        if (IN_OUT == "green") {dots.IN <- dots.IN +1}
    }
    
#     ## Triagle area with approximation with 100 data points (gran = 10)
#     AreaApproxGran10 <- round((dots.IN * (max(x) * max(y)) / dim(grid)[1]), 2)

    # create coordinates of grid points 
    gran2 <- 100
    maxx <- max(x); maxy <- max(y)
    scaleX <- maxx/gran2 
    scaleY <- maxy/gran2 
    pX <- seq(0, maxx, scaleX);
    pY <- seq(0, maxy, scaleY);

    grid2 <- data.frame(x = as.numeric(), y = as.numeric())
    for (i in (1:length(pX))) {
        for (j in (1:length(pY))) {
            grid20 <- c(x = pX[i], y = pY[j])
            grid2 <- rbind.data.frame(grid2, grid20)
        }
    }
    names(grid2) <- c("x","y")
    
    ## PLot the triangle with the coloured gridpoints
    dots.IN2 <- 0
    for (k in (1:dim(grid2)[1])) {
        side <- NULL
        ## Point inside or outside of polygon?
        for (line in (1:(length(x)-1))) {
            s0 <- ((x[line+1] - x[line]) * (grid2[k,2] - y[line]) 
                   - (y[line+1] - y[line]) * (grid2[k,1] - x[line]))
            side <- c(side, s0)
        }
        IN_OUT <- ifelse(sum(abs(side)) == abs(sum(side))
                         , "green", "blue") # "in", "out"
        if (IN_OUT == "green") {dots.IN2 <- dots.IN2 +1}
    }
```

#### While calculation with Heron's formula gives the arithmetically correct area of the triagle the geometric approximation would only give the same results with infinite number of grid points. 
#### We can see that with more grid points the area calculated with approximated approximation is getting closer to the real value.

#### A triangle with coordinates: `r triangle`
```{r}
    # previous calculations are hidden to conserve space
    ## Triagle area with Heron's formula
    AreaHeron <- round((abs(x[1]*y[2] + x[2]*y[3] + x[3]*y[1] - x[1]*y[3] - x[2]*y[1] - x[3]*y[2]) / 2), 2)
    ## Triagle area with approximation with 100 data points (gran = 10)
    AreaApproxGran10 <- round((dots.IN * (max(x) * max(y)) / dim(grid)[1]), 2)
    ## Triagle area with approximation with 1000 data points (gran = 100)
    AreaApproxGran100 <- round((dots.IN2 * (max(x) * max(y)) / dim(grid2)[1]), 2)
```

#### Area with Heron's formula: __`r I(AreaHeron)`__ squared units

#### Area with approximation, granularity = 10 (100 grid points): __`r I(AreaApproxGran10)`__ squared units

#### Area with approximation, granularity = 100 (1000 grid points): __`r I(AreaApproxGran100)`__ squared units



